---
title: "Simulation Error"
output: html_document
---


Simulation generates approximate answers; there is some degree of error in a quantity estimated by Monte Carlo simulation. Intuitively, it seems that the degree of error should get smaller as the number of simulation replicates increases. In this assignment, you are going to investigate the relationship between then number of replicates and simulation error.


## Basic Function 
p̂ is the probbility estimated from simulation

p is the true underlying probability 

absolute error=|p̂−p|

relative error=|p̂−p|/p.



## 14 X 5 factorial experiment simulation
The simulation is based on replicate number (22, 23, …, 215) and probability (0.01, 0.05, 0.10, 0.25, 0.50)
```{r}
library(tidyverse)
library(ggplot2)
output <- expand_grid(
  N= 2^c(2:15)
, P= c(0.01, 0.05, 0.10, 0.25, 0.50)
, abs_error= NA
, rel_error= NA
, KEEP.OUT.ATTRS = FALSE

)
#output <- arrange(output, P)
r <- 10000
for (i in 1:nrow(output)) {
  p <- output$P[i]
  n <- output$N[i]
  phat <- rbinom(r,n,p)/n
  output[i,"abs_error"] <-mean(abs(phat - p))
  output[i,"rel_error"] <-mean(abs(phat - p)/p)
}






```


## Figure: Absolute Error
```{r}
ggplot(output,aes( N, abs_error) )+ geom_point(aes( color=P),size = 2) + geom_line(aes(group = P,color=P),size = 1) + scale_x_continuous(trans = "log2") + xlab("N (log2 scale)")  
```



## Figure: Relative Error 
```{r}
ggplot(output,aes( N, rel_error) )+ geom_point(aes(color=P),size = 2) + geom_line(aes(group = P,color=P),size = 1) + scale_x_continuous(trans = "log2") + xlab("N (log2 scale)") 

```

## Figure: Absolute Error with log10 scale on y-axis

```{r}

ggplot(output,aes( N, log10(abs_error)) )+ geom_point(aes(color=P),size = 2) + geom_line(aes(group = P,color=P),size = 1) + scale_x_continuous(trans = "log2") + xlab("N (log2 scale)") 
```
The relationship between Log10 abserror and N varies before 32. after 32, it shows a linear relationship. 

## Figure: Relative Error with log10 scale on y-axis

```{r}
ggplot(output,aes( N, log10(rel_error)))+ geom_point(aes(color=P),size = 2) + geom_line(aes(group = P,color=P),size = 1) + scale_x_continuous(trans = "log2") + xlab("N (log2 scale)") 
```
The relationship between Log10 abserror and N varies before 32. after 32, it shows a linear relationship. 

## Conclusion 
Our prediction was proved. As the number of replicates increases, the degree of error reduced. We can observe that the N and log10 for both relative error and absolute error have a linear relationship after 32.  