---
title: "05-log-transformation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Gamma Cdf & Pdf
```{r}


x=seq(0.1,20,0.001)
yvalue<-dgamma(x,3,1)
x.mean<-3*1
x.median<- qgamma(0.5,3,1)
plot(x,dgamma(x,3,1),main=
       "Gama Probability Density Function",ylab = "P",xlab = "x",type = "l") + abline(v=x.mean,col="red") + abline(v=x.median,col="blue") 
legend("topright",c("mean","median"),col = c("red","blue"),lty = c(1,1))

```
This is the PDF figure for gamma distribution. 

```{r}
x.mean<-3*1
x.median<- qgamma(0.5,3,1)
plot(x,pgamma(x,3,1),main=
       "Gama Cumulative Distribution Function",ylab = "P",xlab = "x",type = "l") + abline(v=x.mean,col="red",labs="m") + abline(v=x.median,col="blue")
legend("topright",c("mean","median"),col = c("red","blue"),lty = c(1,1))


```
This is the CDF figure for gamma distribution. 


```{r}
x <- rgamma(10000,3,1)
x.mean<-mean(log(x))
x.median<- median(log(x))
plot(density(log(x)),type = "l",main=
       "Gama Log Transformation Probability Density Function",ylab = "P",xlab = "x") + abline(v=x.mean,col="red")+abline(v=x.median,col="blue")
legend("topright",c("mean","median"),col = c("red","blue"),lty = c(1,1))
```
This is the PDF figure for log gamma distribution. 


```{r}

plot(ecdf(log(x)), main= "Gama Log Transformation Cumulative Distribution Function") + abline(v=x.mean,col="red",labs="m") + abline(v=x.median,col="blue")
legend("topright",c("mean","median"),col = c("red","blue"),lty = c(1,1))

#text(x.mean,"mean")
#text(x.median,"median")
```
This is the CDF figure for log gamma distribution. 

## Log Normal Cdf & Pdf
```{r}
x=seq(0,20,0.001)
x.mean<-exp(-1 + 1/2)
x.median<- exp(-1)
plot(x,dlnorm(x,-1,1),main=
       "Log Normal Probability Density Function",ylab = "P",xlab = "x",type = "l") + abline(v=x.mean,col="red",labs="m") + abline(v=x.median,col="blue") 
```
This is the PDF figure for Nomral distribution. 


```{r}
plot(x,plnorm(x,-1,1),main=
       "Log Normal Cumulative Distribution Function",ylab = "P",type="l",xlab = "x") + abline(v=x.mean,col="red",labs="m") + abline(v=x.median,col="blue")

#x=seq(0,20,0.001)
```
This is the CDF figure for Nomral distribution. 


```{r}
x.mean<-mean(log(rlnorm(1000,-1,1)))
x.median<- median(log(rlnorm(1000,-1,1)))
plot(density(log(rlnorm(1000,-1,1))),main=
       "Log Normal Log Transformation Probability Density Function",type = "l") + abline(v=x.mean,col="red",labs="m") + abline(v=x.median,col="blue") 
```
This is the PDF figure for log Nomral distribution. 


```{r}
plot(ecdf(log(rlnorm(1000,-1,1))),main=
       "Log Normal Log Transformation Cumulative Distribution Function",ylab = "P",xlab = "x") + abline(v=x.mean,col="red",labs="m") + abline(v=x.median,col="blue")
#plot(x,plnorm(x,-1,1))
```
This is the CDF figure for log Nomral distribution. 



## Uniform Cdf & Pdf
```{r}
x=seq(0,20,0.001)
x.mean<-(0+12)/2
x.median<- (0+12)/2
plot(x,dunif(x,0,12),main=
       "Uniform Probability Density Function",ylab = "P",xlab = "x",type = "l") + abline(v=x.mean,col="red",labs="m") + abline(v=x.median,col="blue") 
legend("topright",c("mean","median"),col = c("red","blue"),lty = c(1,1))
```
This is the PDF figure for uniform distribution. 


```{r}
plot(x,punif(x,0,12),main=
       "Uniform Cumulative Distribution Function",ylab = "P",type="l",xlab = "x") + abline(v=x.mean,col="red",labs="m") + abline(v=x.median,col="blue")
legend("topright",c("mean","median"),col = c("red","blue"),lty = c(1,1))
```


```{r}
x=seq(0,20,0.001)
x.mean<-mean(log(runif(1000,0,12)))
x.median<- median(log(runif(1000,0,12)))
plot(density(log(runif(1000,0,12))),main=
       "Uniform Log Transformation Probability Density Function") + abline(v=x.mean,col="red") + abline(v=x.median,col="blue") 
legend("topright",c("mean","median"),col = c("red","blue"),lty = c(1,1))

```


```{r}
plot(ecdf(log(runif(1000,0,12))),main=
       "Uniform Log Transformation Cumulative Distribution Function",ylab = "P",xlab = "x") + abline(v=x.mean,col="red") + abline(v=x.median,col="blue")
legend("topright",c("mean","median"),col = c("red","blue"),lty = c(1,1))

```
```{r}
a<-rep(NA,1000)
b<-rep(NA,1000)
for (i in 1:1000) {
        c=rgamma(100,3,1)
        a[i]=mean(c)
        b[i]=exp(mean(log(c)))
        
}
plot(a,b,main="geometic and arithmetic mean",xlab="Arithmetic Mean",ylab="Geometic Mean")
abline(lm(b ~ a))
hist(x=(a-b),main="Difference between Geometic and Arithmetic mean",xlab="Difference")

```

```{r}
a<-rep(NA,1000)
b<-rep(NA,1000)
for (i in 1:1000) {
        c=rlnorm(100,-1,1)
        a[i]=mean(c)
        b[i]=exp(mean(log(c)))
        
}
plot(a,b,main="geometic and arithmetic mean",xlab="Arithmetic Mean",ylab="Geometic Mean")
abline(lm(b ~ a))
hist(x=(a-b),main="Difference between Geometic and Arithmetic mean",xlab="Difference")

```



```{r}
a<-rep(NA,1000)
b<-rep(NA,1000)
for (i in 1:1000) {
        c=runif(100,0,12)
        a[i]=mean(c)
        b[i]=exp(mean(log(c)))
        
}
plot(a,b,main="geometic and arithmetic mean",xlab="Arithmetic Mean",ylab="Geometic Mean")
abline(lm(b ~ a))
hist(x=(a-b),main="Difference between Geometic and Arithmetic mean",xlab="Difference")

```



```{r}
x=0:10
exp(log(x))
log(exp(x))
```

```{r}
```


```{r}
```

