---
title: "05-log-transformation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## For each distribution below, generate a figure of the PDF and CDF. Mark the mean and median in the figure.

## For each distribution below, generate a figure of the PDF and CDF of the transformation Y = log(X) random variable. Mark the mean and median in the figure. You may use simulation or analytic methods in order find the PDF and CDF of the transformation.

##  For each of the distributions below, generate 1000 samples of size 100. For each sample, calculate the geometric and arithmetic mean. Generate a scatter plot of the geometic and arithmetic sample means. Add the line of identify as a reference line.

## Generate a histogram of the difference between the arithmetic mean and the geometric mean.




## Gamma Cdf & Pdf
```{r}


x=seq(0.1,20,0.001)
yvalue<-dgamma(x,3,1)
x.mean<-3*1
x.median<- qgamma(0.5,3,1)
plot(x,dgamma(x,3,1),main=
       "Gamma Probability Density Function",ylab = "P",xlab = "x",type = "l") + abline(v=x.mean,col="red") + abline(v=x.median,col="blue") 
legend("topright",c("mean","median"),col = c("red","blue"),lty = c(1,1))

```
This is the PDF figure for gamma distribution. 

```{r}
x.mean<-3*1
x.median<- qgamma(0.5,3,1)
plot(x,pgamma(x,3,1),main=
       "Gama Cumulative Distribution Function",ylab = "P",xlab = "x",type = "l") + abline(v=x.mean,col="red",labs="m") + abline(v=x.median,col="blue")
legend("topright",c("mean","median"),col = c("red","blue"),lty = c(1,1))


```
This is the CDF figure for gamma distribution. 


```{r}
x <- rgamma(10000,3,1)
x.mean<-mean(log(x))
x.median<- median(log(x))
plot(density(log(x)),type = "l",main=
       "Gama Log Transformation Probability Density Function",ylab = "P",xlab = "x") + abline(v=x.mean,col="red")+abline(v=x.median,col="blue")
legend("topright",c("mean","median"),col = c("red","blue"),lty = c(1,1))
```
This is the PDF figure for log gamma distribution. 


```{r}

plot(ecdf(log(x)), main= "Gama Log Transformation Cumulative Distribution Function") + abline(v=x.mean,col="red",labs="m") + abline(v=x.median,col="blue")
legend("topright",c("mean","median"),col = c("red","blue"),lty = c(1,1))

#text(x.mean,"mean")
#text(x.median,"median")
```
This is the CDF figure for log gamma distribution. 

## Log Normal Cdf & Pdf
```{r}
x=seq(0,20,0.001)
x.mean<-exp(-1 + 1/2)
x.median<- exp(-1)
plot(x,dlnorm(x,-1,1),main=
       "Log Normal Probability Density Function",ylab = "P",xlab = "x",type = "l") + abline(v=x.mean,col="red",labs="m") + abline(v=x.median,col="blue") 
```
This is the PDF figure for Nomral distribution. 


```{r}
plot(x,plnorm(x,-1,1),main=
       "Log Normal Cumulative Distribution Function",ylab = "P",type="l",xlab = "x") + abline(v=x.mean,col="red",labs="m") + abline(v=x.median,col="blue")

#x=seq(0,20,0.001)
```
This is the CDF figure for Nomral distribution. 


```{r}
x.mean<-mean(log(rlnorm(1000,-1,1)))
x.median<- median(log(rlnorm(1000,-1,1)))
plot(density(log(rlnorm(1000,-1,1))),main=
       "Log Normal Log Transformation Probability Density Function",type = "l") + abline(v=x.mean,col="red",labs="m") + abline(v=x.median,col="blue") 
```
This is the PDF figure for log Nomral distribution. 


```{r}
plot(ecdf(log(rlnorm(1000,-1,1))),main=
       "Log Normal Log Transformation Cumulative Distribution Function",ylab = "P",xlab = "x") + abline(v=x.mean,col="red",labs="m") + abline(v=x.median,col="blue")
#plot(x,plnorm(x,-1,1))
```
This is the CDF figure for log Nomral distribution. 



## Uniform Cdf & Pdf
```{r}
x=seq(0,20,0.001)
x.mean<-(0+12)/2
x.median<- (0+12)/2
plot(x,dunif(x,0,12),main=
       "Uniform Probability Density Function",ylab = "P",xlab = "x",type = "l") + abline(v=x.mean,col="red",labs="m") + abline(v=x.median,col="blue") 
legend("topright",c("mean","median"),col = c("red","blue"),lty = c(1,1))
```
This is the PDF figure for uniform distribution. 


```{r}
plot(x,punif(x,0,12),main=
       "Uniform Cumulative Distribution Function",ylab = "P",type="l",xlab = "x") + abline(v=x.mean,col="red",labs="m") + abline(v=x.median,col="blue")
legend("topright",c("mean","median"),col = c("red","blue"),lty = c(1,1))
```
This is the CDF for uniform distribution 

```{r}
x=seq(0,20,0.001)
x.mean<-mean(log(runif(1000,0,12)))
x.median<- median(log(runif(1000,0,12)))
plot(density(log(runif(1000,0,12))),main=
       "Uniform Log Transformation Probability Density Function") + abline(v=x.mean,col="red") + abline(v=x.median,col="blue") 
legend("topright",c("mean","median"),col = c("red","blue"),lty = c(1,1))

```
This is the PDF for log uniform distribution.

```{r}
plot(ecdf(log(runif(1000,0,12))),main=
       "Uniform Log Transformation Cumulative Distribution Function",ylab = "P",xlab = "x") + abline(v=x.mean,col="red") + abline(v=x.median,col="blue")
legend("topright",c("mean","median"),col = c("red","blue"),lty = c(1,1))

```

This the CDF for log uniform distribution. 

```{r}
a<-rep(NA,1000)
b<-rep(NA,1000)
for (i in 1:1000) {
        c=rgamma(100,3,1)
        a[i]=mean(c)
        b[i]=exp(mean(log(c)))
        
}
plot(a,b,main="geometic and arithmetic mean for gamma distribution",xlab="Arithmetic Mean",ylab="Geometic Mean")
abline(0,1)
hist(x=(a-b),main="Difference between Geometic and Arithmetic mean for gamma distribution ",xlab="Difference")

```
In this section, we used a loop to simulate the situation. It has a 1000 samples of size 100. Both the geometric and arithmetic mean is ploted in the same graph. The X is arithmetic mean and the y is geometic mean. The histogram shows the difference between them for each individual value. 

```{r}
a<-rep(NA,1000)
b<-rep(NA,1000)
for (i in 1:1000) {
        c=rlnorm(100,-1,1)
        a[i]=mean(c)
        b[i]=exp(mean(log(c)))
        
}
plot(a,b,main="geometic and arithmetic mean for log normal distribution",xlab="Arithmetic Mean",ylab="Geometic Mean")
abline(0,1)
hist(x=(a-b),main="Difference between Geometic and Arithmetic mean for log normal distribution",xlab="Difference")

```

In this section, we used a loop to simulate the situation. It has a 1000 samples of size 100. Both the geometric and arithmetic mean is ploted in the same graph. The X is arithmetic mean and the y is geometic mean. The histogram shows the difference between them for each individual value. 


```{r}
a<-rep(NA,1000)
b<-rep(NA,1000)
for (i in 1:1000) {
        c=runif(100,0,12)
        a[i]=mean(c)
        b[i]=exp(mean(log(c)))
        
}
plot(a,b,main="geometic and arithmetic mean for uniform distribution ",xlab="Arithmetic Mean",ylab="Geometic Mean")
abline(0,1)
hist(x=(a-b),main="Difference between Geometic and Arithmetic mean for uniform distribution",xlab="Difference")

```
In this section, we used a loop to simulate the situation. It has a 1000 samples of size 100. Both the geometric and arithmetic mean is ploted in the same graph. The X is arithmetic mean and the y is geometic mean. The histogram shows the difference between them for each individual value. 


## part 3 

```{r}
relationship <- function(sample,size){
  a<-rep(NA,size)
  b<-rep(NA,size)
  diff<-rep(NA,size)
  c<-NA
  for (i in 1:size) {
    c=sample(sample,size,replace = TRUE)
    a[i]=log(mean(c))
    b[i]=mean(log(c))
    diff[i]=a[i]-b[i]
  }
  #mean(diff)
  diff
  }
relationship(1000,100)


d<-rep(1,100)
e=log(mean(d))
f=mean(log(d))
e == f


```
By running this simulation, we can see that they are not equal. By looking at the result, all of the simulations are positive, the log (E[X]) is greater than E[log (X)]. This simulation has ran for a 100 samples of size 1000. The diff is the difference between those two value. For a special condition that all of the values are 1, both the log (E[X]) and E[log (X)] will return 0, which means they are equal. For conclusion, log (E[X]) is laways greater or equal to E[log (X)].



